---
title: "Scrape Wikipedia_COVID-19_pandemic_lockdowns"
output: html_document
author: Glen Haynes
params:
  use_saved_files: TRUE
  use_wiki_csv: FALSE
---

Setup

```{r}
# install.packages('tidyverse')
# install.packages('tidyselect')
# install.packages('purrr')
# install.packages('jsonlite')
# install.packages('DT')
# install.packages('tibble')
# install.packages('reticulate')
# install.packages('rvest')
# install.packages('robotstxt')
# install.packages('readxl')
# install.packages('datasets')
# install.packages('growthrates')
# install.packages('readr')
# install.packages('ggcorrplot')
# install.packages('GGally')
```

```{r}
library(tidyverse)
library(tidyselect)
library(purrr)
library(jsonlite)
library(DT)
library(tibble)
library(reticulate)
library(rvest)
library(robotstxt)
library(readxl)
library(datasets)
library(growthrates)
library(readr)
library(ggcorrplot)
library(GGally)
```

# Data Collection

## Johns Hopkins CSSE Data

```{r}

## Load functions ##

source("R/csse_covid_dat.R")
source("R/csse_merge.R")

## List CSSE files ##

csse_list <- list(
  "time_series_covid19_confirmed_US.csv",
  "time_series_covid19_confirmed_global.csv",
  "time_series_covid19_deaths_US.csv",
  "time_series_covid19_deaths_global.csv",
  "time_series_covid19_recovered_global.csv"
)

csse_dat <- csse_list %>% map(~{
  csse_covid_dat(.x, 
                 use_saved_file = params$use_saved_files)
})

csse_us_merge <- csse_merge(csse_dat, 'us', 'Series')

csse_global_merge <- csse_merge(csse_dat, 'global', 'Series')

rm(csse_dat)
```

## World Bank Data

```{r}
source("R/wrld_bk_dat.R")
source("R/wb_ind_lst.R")

## Get World Bank meta data ##

wb_meta <- wrld_bk_dat(use_local_json = params$use_saved_files)

## List of relevant indicators ##

if(params$use_saved_files)
{
  wb_relevant_indicators <- wb_ind_lst(dat_loc = "data/covid-indicators.csv")
}else
{
  wb_relevant_indicators <- wb_ind_lst()
}

## Limit relevant indicators ##

wb_relevant_indicators <- wb_relevant_indicators[c(1, 5, 18, 25, 37, 47),]

## Relevant indicators and other data ##

wb_rel_ind_List <- map2(
  c(wb_relevant_indicators$id, 'SP.POP.TOTL'), ## Add Population
  c(wb_relevant_indicators$topic, "Extra_to_rel"),
  function(x,y){wrld_bk_dat(indicator = x,
                            date='2015:2020',
                            use_local_json = TRUE,
                            save_file = FALSE,
                            save_dir = 'data/wb-rel-ind-data',
                            dat_addrs = NULL) %>%
      mutate(topic = y) %>% 
      group_by(countryiso3code) %>%
      group_map(~ .x %>% 
                  filter(!is.na(value)) %>%
                  filter(date == max(date))) %>% 
      bind_rows()}) %>% 
  bind_rows()

```

## Wikipedia Lockdown Data

# Python Code was adapted from https://simpleanalytical.com/how-to-web-scrape-wikipedia-python-urllib-beautiful-soup-pandas #

```{r}
if(params$use_wiki_csv)
{
  wiki_lock <- read_csv(wiki_lock, "data/wiki_lock.csv")
  wiki_no_lock<- read_csv(wiki_no_lock, "data/wiki_no_lock.csv")
}else
{
  if(params$use_saved_files & !params$use_wiki_csv)
  {
    use_python("c/Miniconda3/python")
    wiki_covid_lockdown <- import("wiki_covid_lockdown")
    
    wiki_lock_status <- wiki_covid_lockdown$wiki_covid_lockdown(
      source = "web_page_references/COVID-19 pandemic lockdowns - Wikipedia.html"
    )
    
    wiki_lock <- wiki_lock_status[[1]] %>%
      map(~sapply(.x, function(y){
        as.character(y) %>% gsub("^\\s+|\\s+$", "", .)
      })) %>% bind_cols()
    
    wiki_no_lock <- wiki_lock_status[[2]] %>%
      map(~sapply(.x, function(y){
        as.character(y) %>% gsub("^\\s+|\\s+$", "", .)
      })) %>% bind_cols()
  }else
  {
    
    ## Check Wikipeida still allowing bots ##
    
    wiki_rbt_status <- robotstxt(domain="en.wikipedia.org")
    
    # Check access permission (returns true/false)
    
    wiki_scrape_allowed <- wiki_rbt_status$check(
      paths = "wiki/COVID-19_pandemic_lockdowns", 
      bot = "*")
    
    if(wiki_scrape_allowed)
    {
      
      use_python("c/Miniconda3/python")
      wiki_covid_lockdown <- import("wiki_covid_lockdown")
      
      wiki_lock_status <- wiki_covid_lockdown$wiki_covid_lockdown()
      
    }
    
    wiki_lock <- wiki_lock_status[[1]] %>%
      map(~sapply(.x, function(y){
        as.character(y) %>% gsub("^\\s+|\\s+$", "", .)
      })) %>% bind_cols() %>% 
      ## Convert start and end dates to R date type ##
      mutate(`Start date` = as.Date(`Start date`, format = "%Y-%m-%d")) %>%
      mutate(`End date` = as.Date(`End date`, format = "%Y-%m-%d")) %>% 
      rename(start_date = `Start date`, end_date = `End date`)
    
    wiki_no_lock <- wiki_lock_status[[2]] %>%
      map(~sapply(.x, function(y){
        as.character(y) %>% gsub("^\\s+|\\s+$", "", .)
      })) %>% bind_cols()
  }
}

```

# Littler US State Level Lockdowns

```{r}
source("R/littler_us_st_sty_hm.R")

litt_state <- littler_us_st_sty_hm(url = paste0("https://www.littler.com/",
                                  "publication-press/publication/",
                                  "stay-top-stay-home-list",
                                  "-statewide"),
                     save_file = TRUE,
                     save_dir = 'data',
                     use_local_file = FALSE,
                     add_na_states = TRUE
)
```

# Create crosswalks

```{r}
source("R/csse_wb_crs_wlk.R")
source("R/wiki_crs_wlk.R")

## Countries ##

global_csse_crs <- csse_global_merge %>% 
  csse_wb_crs_wlk(wb_meta) %>% 
  wiki_crs_wlk(wiki_lock) %>% 
  filter(!is.na(csse_cntry))


## U.S. States and D.C. ##

# Combine list U.S. states list (from core package 'datasets') with # 
# "District of Columbia" #

state_dc <- c(datasets::state.name)

# U.S. territories ##

terr <- c("Puerto Rico",
          "American Samoa",
          "Guam",
          "Northern Mariana Islands",
          "Virgin Islands")

```

# CSSE Data Transfomrations

## Filter for number of events less than 1000, get totals, and calculate growth rates. Filter to exclude sub-country regions.

```{r, warning = FALSE}
## Filter out Non-Occurances across Confirmed Cases, Deaths and Recovered ##

source("R/csse_to_long.R")
source("R/simple_avg_growth.R")
source("R/cmpnd_avg_growth.R")
source("R/implement_fit_easylinear.R")

csse_events <- csse_global_merge %>% 
  filter(., {select(., matches("\\d+/\\d+/\\d+")) %>% 
      apply(1, function(x){!all(x < 1000)})}) %>% 
  filter(is.na(`Province/State`)) %>% 
  select(-`Province/State`)

## Calculate Growth Rates (Simple Average, Compound Average, and Maximum)

csse_grwth <- map(list("confirmed", "deaths", "recovered"), function(srs)
{
  grw1 <- csse_events %>%
    csse_to_long(srs) %>% 
    group_by_at(vars(-one_of("Series", "Date", "value"))) %>% 
    summarise(
      smpl_growth = simple_avg_growth(value, start_val = 10),
      cmpnd_growth = cmpnd_avg_growth(value, start_val = 10),
      start = if(any(value >= 10))
        {Date[(which(value >= 10))[1]]}else{as.Date("9999-01-01")},
      max = max(value),
      t = sum(value >= 10)
    ) 
  
  grw2 <- csse_events %>%
    csse_to_long(srs) %>% 
    group_by_at(vars(-one_of("Series", "Date", "value"))) %>%
    summarise(max_grw = implement_fit_easylinear(value)[[1]],
              max_grw_r2 = implement_fit_easylinear(value)[[2]])
  
  grw_rtrn <- full_join(grw1, grw2, by = intersect(colnames(grw1), 
                                                   colnames(grw2))) %>% 
    mutate(Series = srs) %>% 
 select(-max, -t, -Series, everything())
}) %>% 
  bind_rows()

```

# Sort and plot CSSE Data

```{r, fig.height=10, fig.width=10, warning=FALSE,echo=FALSE}
sort_csse <- lapply(unlist(csse_grwth[grepl("deaths|recovered", 
                                                csse_grwth$Series),
                       "Country/Region"]), function(x)
  {
    csse_grwth %>%
  ungroup() %>% 
  filter(`Country/Region` == x) %>% 
  filter(grepl("deaths|recovered", Series)) %>%
  select(-one_of(c("Lat", "Long", "max_grw_r2"))) %>%
  group_by(`Country/Region`, Series, start) %>%
  group_map(~
              if(nrow(.y[,2]))
              {
                .y[,2] <- paste0(.y[,2], "_global")
                    left_join(.y[,-3], 
                              csse_events[grepl("deaths|recovered", 
                                                csse_events$Series),],
                                    by = intersect(colnames(csse_events),
                                                   names(.y))
                                ) %>%
                      csse_to_long("*") %>%
                        filter(Date >= max(.y$start)) %>%
                      mutate(smpl_growth = .x$smpl_growth * 1:nrow(.)) %>%
                      mutate(cmpnd_growth = .x$cmpnd_growth * 1:nrow(.)) %>%
                      mutate(max_grw = .x$max_grw * 1:nrow(.))
              }) %>%
  bind_rows() %>%
  select(-one_of(c("Lat", "Long"))) %>%
  rename(raw = value, country = `Country/Region`) %>%
                           left_join(global_csse_crs, 
                                     by = c("country" = "csse_cntry")) %>% 
                           left_join({wiki_lock %>% 
  group_by(`Countries and territories`) %>% 
  summarise(mx_start = max(`Start date`))}, 
                                     by = c("wiki_country" = 
                                              "Countries and territories")) %>% 
  pivot_longer(one_of(c("raw", "smpl_growth", "cmpnd_growth", "max_grw")),
             names_to = "mthd", values_to = "n") %>%
                           mutate(mx_start = as.Date(mx_start)) %>%
                           mutate(Series = gsub("global", x, Series))
                       })
                         
plts <- lapply(sort_csse, function(x)
         {x %>% 
  {
    mx_strt <- unique(.$mx_start)
    if(!is.na(mx_strt))
    {
      v_line <- geom_vline(xintercept = mx_strt)
      v_line2 <- geom_vline(xintercept = (mx_strt + 31), linetype = "dashed")
    }else
    {
      v_line <- NULL
      v_line2 <- NULL
    }
    ggplot(data = .) +
      geom_line(mapping = aes(x = Date, y = n, color = mthd)) +
      v_line +
      v_line2 +
      facet_wrap(vars(Series), ncol = 2, nrow = 2)
  }

  })

# pdf("plots/raw_against_growth.pdf")
# plts
# dev.off()

```

```{r}
source("R/join_glob_dat.R")
csse_joined <- join_glob_dat(csse_dat = csse_grwth,
                             lockdown_dat = wiki_lock,
                             wb_dat = wb_rel_ind_List,
                             wb_met = NA,
                             crss_walk = global_csse_crs)
 
```

# Examine correlations from World Bank's list relevant indicators

```{r}
source("R/fltr_crss_wb_cat.R")
wb_dat <- csse_joined %>%
  ungroup %>% 
  select(topic, indicator.id, wb_name, value) %>%
  mutate(value = as.numeric(value)) %>% 
  unique() %>% 
  left_join(select(wb_meta, name, region.value, 
                                incomeLevel.value), by = c("wb_name" ="name"))

corr_dat <- lapply(unique(wb_meta$incomeLevel.value), function(x)
  {
  inc_rgn <- lapply(unique(wb_meta$region.value), function(y)
  {
    corr_wb <- wb_dat %>% 
      filter(wb_name %in% {fltr_crss_wb_cat(incomeLevel.value, x) %>% 
               fltr_crss_wb_cat(region.value, y, .)}$wb_name) %>% 
      select(-topic, -incomeLevel.value, -region.value) %>%
      data.frame(stringsAsFactors = FALSE) %>% 
      pivot_wider(names_from = indicator.id, values_from = value) %>%
      
      {if(nrow(.))
      {
        tmp_dat <- data.frame(., stringsAsFactors = FALSE)
        rownames(tmp_dat) <- tmp_dat$indicator.value
        tmp_dat <- tmp_dat[,-1]
        tmp_dat
      }else
      {
        NULL
      }
      }
  })
  names(inc_rgn) <- paste(unique(wb_meta$region.value), x, sep = ": ")
  return(inc_rgn)
}) %>% lapply(., function(x){Filter(Negate(is.null),x)}) %>% 
  unlist(., recursive = FALSE)

corr_rslt <- corr_dat  %>% 
  map(~cor(.x, method = "pearson", 
                  use = "pairwise.complete.obs")) %>% 
  Filter(function(x){!all(is.na(x)|abs(x) == 1)}, .)

corr_dat <- corr_dat[names(corr_dat) %in% names(corr_rslt)]
  
corr_plts <- map2(corr_dat, names(corr_dat), function(x, y)
{
  ggcorr(x, label = TRUE) +
  ggtitle(y)
})



pdf("plots/corr_plots.pdf")
corr_plts
dev.off()

```


# Filter CSSE data with world bank categories and link with Wikipedia lockdown information
## Take countries with no lockdown order and group them by their 'region.value' and 'incomeLevel.value from the World Bank.
## Format the sorted data by creating before and after lockdown dummy and dummy for in-or-out of lockdown set. (Actually take 31 days after lockdown imposed as binary turning point for the time dummy accounting for the lag effect in mortality and recovery.)


```{r}
source("R/fltr_crss_wb_cat.R")
no_lock_countries <- c(wiki_no_lock$`Countries and territories`[
  wiki_no_lock$`Countries and territories` %in% 
    global_csse_crs$wb_name][-5], "Korea, Rep.")

csse_no_lock <- no_lock_countries %>% map(~
                            {
                              wb_meta %>% filter(
                                {
                                  name %in% .x 
                                }
                              )%>%  
                                {
                                  left_join(select(., 
                                                   incomeLevel.value, 
                                                   region.value),
                                             wb_meta, 
                                            by = c("incomeLevel.value", 
                                                   "region.value")) %>%
                                    select(name)
                                } %>%
                                unlist() %>% 
                                paste(collapse = "|") %>% 
                                fltr_crss_wb_cat(name, .) %>% 
                                select(csse_cntry) %>%
                                { 
                                  filter(bind_rows(sort_csse),
                                         country %in% unlist(.)) %>%
                                    #### Create treatment effect dummy  ##
                                    ## Filter out all non-locked except ##
                                    ## non-locked of interest in list the  ##
                                    ## element. Later each locked country ##
                                    ## will be run in a seperate equation ##
                                    ## against the locked in it's group/list. ##
                                    mutate(lock = ifelse(wb_name %in%
                                                           no_lock_countries,
                                      0, 1)) %>% 
                                    filter(!(lock == "N" & wb_name == .x)) %>% 
                                    mutate(Start_Lag = mx_start + 31) %>% 
                                    filter(mthd == "raw" & grepl("death|recov",
                                                                 Series)) %>% 
                                    select(country, Series, 
                                           Date, Start_Lag, n, lock)
                                }
                            })
```

## Finally model difference-in-difference equations and report findings.  
# Differnce in difference for lockdown effect on mortality.

```{r}
diff_in_diff_mortality <- csse_no_lock %>% map2(., no_lock_countries, 
                                                function(x, y)
{
  no_lock_geo <- filter(x, country %in% y)
  lock_geo <- filter(x, !country %in% y)
  lock_geo %>% group_by(country, Series) %>%
    group_map(~
                {
                  model_dat <- cbind(.y, .x) %>% 
                      bind_rows(no_lock_geo) %>% 
                    filter(grepl("death", Series))
                  
                  if(length(unique(model_dat$country)) != 2)
                  {
                    return(NULL)
                  }else
                  {
                    model_dat <- model_dat %>%
                      #### Create time dummy ####
                    ## Use max(na.rm) because unique value and want to extend ##
                    ## the effect to both series. ##
                    mutate(time = ifelse(max(Start_Lag,
                                             na.rm = TRUE) >= Date, 1, 0)) %>%
                      #### Create interaction for Difference in Differences ####
                    mutate(diff_in_diff = time * lock)
                    
                    fit <- tryCatch({lm(n ~ time + lock + diff_in_diff,
                                        data = model_dat)}, error = function(e){
                                          return(NULL)
                                        })
                    return(list(model_dat, fit))
                  }
                  

                })
}) %>% unlist(recursive = FALSE) %>% Filter(Negate(is.null), .)


## Get names from data sets returned ##

names(diff_in_diff_mortality) <- sapply(
  1:length(diff_in_diff_mortality), 
  function(x){diff_in_diff_mortality[[x]][[1]] %>% 
      select(country, lock) %>% 
      unique() %>% apply(., 1, function(x)
      {paste(x, collapse = ":")}) %>% paste(collapse = "-")})

map2(diff_in_diff_mortality, names(diff_in_diff_mortality), function(x, y)
  {
  print(y)
  sum(x[[2]])
  })
```

# Differnce in difference for lockdown effect on recovered.

```{r}
diff_in_diff_recovered <- csse_no_lock %>% map2(., no_lock_countries, 
                                                function(x, y)
{
  no_lock_geo <- filter(x, country %in% y)
  lock_geo <- filter(x, !country %in% y)
  lock_geo %>% group_by(country, Series) %>%
    group_map(~
                {
                  model_dat <- cbind(.y, .x) %>% 
                    bind_rows(no_lock_geo) %>% 
                    filter(grepl("recover", Series))
                  
                  if(length(unique(model_dat$country)) != 2)
                  {
                    return(NULL)
                  }else
                  {
                    model_dat <- model_dat %>%
                      #### Create time dummy ####
                    ## Use max(na.rm) because unique value and want to extend ##
                    ## the effect to both series. ##
                    mutate(time = ifelse(max(Start_Lag,
                                             na.rm = TRUE) >= Date, 1, 0)) %>%
                      #### Create interaction for Difference in Differences ####
                    mutate(diff_in_diff = time * lock)
                    
                    fit <- tryCatch({lm(n ~ time + lock + diff_in_diff,
                                        data = model_dat)}, error = function(e){
                                          return(NULL)
                                        })
                    return(list(model_dat, fit))
                  }
                  
                  
                })
}) %>% unlist(recursive = FALSE) %>% Filter(Negate(is.null), .)

## Get names from data sets returned ##

names(diff_in_diff_recovered) <- sapply(1:length(diff_in_diff_recovered), 
       function(x){diff_in_diff_recovered[[x]][[1]] %>% 
           select(country, lock) %>% 
           unique() %>% apply(., 1, function(x)
             {paste(x, collapse = ":")}) %>% paste(collapse = "-")})

map2(diff_in_diff_recovered, names(diff_in_diff_recovered), function(x, y)
  {
  print(y)
  sum(x[[2]])
  })

```


